\documentclass[11pt,a4paper]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Libraries
\usepackage[driver=pdftex,vmargin=2cm,hmargin=2cm]{geometry}
\usepackage[usenames,dvipsnames]{xcolor}
%% \usepackage{amsmath}
\usepackage{array}
\usepackage{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{paralist}
\usepackage{url}
\usepackage{eurosym}
\usepackage{gb4e}
\usepackage{wasysym}   % smiley symbols

\hypersetup{
  colorlinks,
  citecolor=Violet,
  linkcolor=Red,
  urlcolor=Blue}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Commands
\definecolor{dodgerblue4}{RGB}{16,78,139}
\definecolor{orange3}{RGB}{205,133,0}
\definecolor{DarkSlateBlue}{RGB}{72,61,139}

\newcommand{\xmltag}[1]{{\textbf{\small$<$#1$>$}}}
\newcommand{\sentiment}[1]{\xmltag{sentiment}#1\xmltag{/sentiment}}
\newcommand{\source}[1]{\xmltag{source}#1\xmltag{/source}}
\newcommand{\target}[1]{\xmltag{target}#1\xmltag{/target}}
\newcommand{\emoexpression}[1]{\xmltag{emo-expression}#1\xmltag{/emo-expression}}
\newcommand{\intensifier}[1]{\xmltag{intensifier}#1\xmltag{/intensifier}}
\newcommand{\diminisher}[1]{\xmltag{diminisher}#1\xmltag{/diminisher}}
\newcommand{\negation}[1]{\xmltag{negation}#1\xmltag{/negation}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Lengths
\newlength{\cwidth} \setlength{\cwidth}{0.18\textwidth}
\newlength\clmnwidth
\setlength{\clmnwidth}{0.6\textwidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Environments
\newenvironment{myexe}{
  \begin{exe}
    \ex\begin{center}
    \itshape
}{
    \end{center}
  \end{exe}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Title
\author{Wladimir Sidorenko}
\date{\today}
\title{Guidelines for the Annotation of the Sentiment Corpus}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Main
\begin{document}
\maketitle{}
\section{Overview}
\subsection{Introduction}

Welcome.  In this assignment, your task will be to annotate sentiments in a
corpus of German Twitter messages.

For the purpose of this assignment, we will regard as sentiments polar
assessing opinions about some persons, subjects, or events.  Your task will be
to annotate both, text spans denoting the assessments (\textit{sentiments})
and text spans representing the subjects and actions being assessed
(\textit{sentiment targets}). In addition to that, you will also have to
annotate the authors of the opinions (also called \textit{sentiment sources}),
words and phrases that bear some polar assessing meaning on their own
(\textit{emotional expressions}), and elements which might decrease or
increase the lexical sense of emotional expressions or turn their meaning to
the opposite.

A complete list of elements to be annotated along with their description and
a description of their attributes is provided in Section \ref{sec:markables}.
Section \ref{sec:markables} will give you a short summary of these elements
and their attributes. In Section \ref{sec:faq}, you will also find some
examples of difficult and controversial annotation cases and some possible
solutions, that we have suggested for them.

\subsection{Annotation Tool}

For the annotation of text, you should use \texttt{MMAX2} -- a freely
available annotation tool.  You can download this program under the
following link:\\\newline
{\setlength{\parindent}{0pt}\small\url{http://sourceforge.net/projects/mmax2/files/mmax2/mmax2_1.13.003/MMAX2_1.13.003b.zip/download}}\\\newline
After downloading, please unpack the received archive, change to the
newly created directory \texttt{1.13.003/MMAX2} in your terminal shell
and execute the following commands:\\

\texttt{> chmod u+x ./mmax2.sh}

\texttt{> nohup ./mmax2.sh \&} \\\newline
       {\setlength{\parindent}{0pt}An \texttt{MMAX} window should then
         appear on your screen.}

If you have never used \texttt{MMAX2} before, please read the document
\texttt{mmax2quickstart.pdf}, which you can find in the subdirectory
\texttt{MMAX2/Docs} of the downloaded archive.

\subsection{Corpus Files}

You should also have received a copy of corpus files as a tar-gzipped
archive.  Please unpack this archive by using the command:\\\newline
\texttt{> tar -xzf twitter-sentiment.tgz}\\\newline
       {\setlength{\parindent}{0pt} A directory called
         \texttt{mmax-prj} should then appear in your current folder.
         Change to your \texttt{MMAX2} window and click on the menu
         \texttt{File -> Load}.  Select the path to the unpacked
         \texttt{mmax-prj} folder in the displayed popup
         window\footnote{Make sure that the path to \texttt{mmax-prj}
           does not contain any white spaces.  Otherwise,
           \texttt{MMAX2} will fail to load the project file.} and
         select one of the \texttt{*.mmax} files there. The source
         data for the annotation file will then be loaded into your
         \texttt{MMAX2} editor.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Tags and Attributes}\label{sec:markables}
Below, you can find a short list of all tags and their possible
attributes, that will be used in this annotation task:

\begin{multicols}{2}
  \begin{enumerate}
  \item \textit{sentiment}-tag with the attributes:
    \begin{enumerate}
    \item polarity,
    \item intensity,
    \item sarcasm;
    \end{enumerate}
  \item \textit{target}-tag with the attributes:
    \begin{enumerate}
    \item anaph-ref,
    \item sentiment-ref;
    \end{enumerate}
  \item \textit{source}-tag with the attributes:
    \begin{enumerate}
    \item anaph-ref,
    \item sentiment-ref;
    \end{enumerate}
  \item \textit{emo-expression}-tag with the attributes:
    \begin{enumerate}
    \item polarity,
    \item intensity,
    \item sarcasm,
    \item sentiment-ref;
    \end{enumerate}
  \item \textit{intensifier}-tag with the attributes:
    \begin{enumerate}
    \item degree,
    \item emo-expression-ref;
    \end{enumerate}
  \item \textit{diminisher}-tag with the attributes:
    \begin{enumerate}
    \item degree,
    \item emo-expression-ref;
    \end{enumerate}
  \item and the \textit{negation}-tag.
  \end{enumerate}
\end{multicols}
A more detailed description of the meaning of these tags and the values of the
attributes is provided in the next sections.

\subsection{sentiment}\label{sec:sentiment}
With \texttt{sentiment} tags, you should mark polar assessing statements which
are made about some subjects or events.  Typical examples of sentiment
relations are sentences like the following one:
\begin{myexe}
  \sentiment{Ich mag den neuen James Bond Film nicht}.

  (\sentiment{I don't like the new James Bond
    movie}.)\label{ex:sentiment}
\end{myexe}
Sentiment expressions, however, are not only constrained to sentences -- they
also can be longer or shorter. In order to decide whether a
given statement forms a sentiment or not, you should check whether it
satisfies the following three conditions:
\begin{enumerate}
  \item The sentiment statement has to be \textit{subjective}, i.e. it has to
    express or imply someone's opinion and should not be a mere declaration of
    objective facts.  Therefore, cases like \textit{CDU hat die Wahl gewonnen
      (CDU has won the election)} should not be marked as sentiments because
    they do not contain any subjective information;

  \item The sentiment statement has to be \textit{polar}, i.e. it should
    clearly express either a positive or a negative judgement about
    something.  According to this rule, cases like \textit{Meiner Meinung nach
      k\"onnte man den Tisch zuerst raustragen (I think we could bring out the
      table first)} should not be marked as sentiments since they do not
    express any polarity (i.e a positive or negative attitude) even though 
	a subjective opinion is expressed;

  \item The judgement of the statement should always be \textit{related to a
    particular target} which can be any object including persons,
    institutions, properties, events etc.  It means that cases like
    \textit{Ich bin so froh! (I am so happy!)} also should not be marked as
    sentiments (however controversial it sounds) since they do not assess
    anything in particular.
\end{enumerate}

Sentiment tags can usually include:
\begin{itemize}
\item single noun phrases, possibly with their prepositional
  attributes, e.g. \textit{Auf dem Tisch lag \xmltag{sentiment}\\ ein
    langweiliges Buch\xmltag{/sentiment} (There was a
    \sentiment{boring book} on the table)};
\item clauses, e.g. \textit{\sentiment{Ich hasse B\"ucher ohne
  Inhaltsangabe}. (\sentiment{I hate books without table of
  contents})};
\item multiple sentences in cases when these sentences jointly form a
  sentiment relation, e.g.\\\textit{\sentiment{Sie denken, reden,
      riechen, lieben, schmecken, f*cken Plastik. Sie haben\\das so
      gelernt in der Plastik-Werbewelt}.\\ (\sentiment{They think and
      speak about, smell at, love, taste, f*ck plastic.  They
      have\\ learned it so in the advertising world of plastic}.)}.
\end{itemize}

Sentiment tags can accept the following attributes with the following
possible values:
\begin{center}
  \begin{tabular}[t]{|l|c|p{\clmnwidth}|}\hline
    Attribute & Value & Value's Meaning\\\hline
    %%%%%%%%%%%%%%%%%%%

    & \textit{positive} & sentiment
    expresses positive attitude to its respective target,
    e.g. \textit{Es war ein fantastischer Abend (It was a fantastic
      eve)};\\\cline{2-3}

    & \textit{negative\newline(default)} & sentiment expresses
    negative attitude to its respective target, e.g. \textit{Seine
      Schwester ist einfach unausstehlich (His sister is simply
      obnoxious)}\\\cline{2-3}

    \multirow{-3}{*}{polarity} & \textit{comparison} & sentiment
    expresses comparison of two targets with preference given to one
    of them, e.g. \textit{Mir gef\"allt das rote Kleid mehr als das
      blaue (I like the red dress more than the blue one)}\\\hline

    %%%%%%%%%%%%%%%%%%%

    & \textit{weak} & weak positive or
    negative assessments, e.g. \textit{Der Auftritt war mehr oder
      weniger gut (The appearance was more or less good)}\\\cline{2-3}

    & \textit{medium\newline(default)} & sentiments with middle
    emotional expressivity, e.g. \textit{Mir hat das neue Album gut
      gefallen (I enjoyed the new album)}\\\cline{2-3}

    \multirow{-3}{*}{intensity} & \textit{strong} & very emotional
    polar statements, e.g. \textit{Dieses Festival war einfach
      umwerfend!!! (This festival was simply terrific!!!)}\\\hline
    %%%%%%%%%%%%%%%%%%%

    \multirow{2}{*}{sarcasm} & \textit{true} & this polar attitude is
    derisive, i.e. its actual polarity is the opposite of its apparent
    form. (This means that an apparent praise which appears in text is
    in fact meant as a rebuke and vice versa. The actual sense,
    however, can only be inferred on the basis of world knowledge or
    reasoning.)  An example of a sarcastic sentiment is the following
    passage: \textit{Mein J\"ungerer ist in der Pr\"ufung
      durchgefallen.  Gut gemacht! (My junior has failed the exam.
      Well done!)}\\\cline{2-3}

    & \textit{false\newline(default)} & no sarcasm is present -- the
    polar attitude has its literal meaning\\\hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{target}
With \texttt{target} tags, you should mark objects or events which are being
assessed by sentiment expressions.  There should always be at least one target
element for one sentiment relation.

An example of a sentiment target is given in the following sentence:
\begin{myexe}
Mein Bruder ist nicht gerade begeistert von \target{dem neuen Call of
  Duty}.

(My brother is not impressed by \target{the new Call of
  Duty}.)
\end{myexe}

The target element is usually expressed by:
\begin{itemize}
\item single noun phrases, possibly with their prepositional attributes,
  e.g. \textit{Mir wird's schlecht, wenn ich \target{diese Werbung} im
    Fernsehen sehe (I feel sick when I see this \target{ad} on TV)};

\item clauses, e.g. \textit{Ich hasse, \target{wenn jemand sich in die dritte
    Spur einordnet und dann wie eine Schnecke kriecht} (I hate \target{when
    someone changes to the leftmost lane and then crawls at a snail's pace})};
\end{itemize}

If the given sentiment relation has multiple targets, which are independent
from each other and conjoined by commas or coordinative conjunctions, you
should mark each of such targets with a separate \texttt{target} tag:
\begin{myexe}
  Meiner Mutter haben immer \target{Nelken} und \target{Dahlien}
  gefallen.

  (My mother has always liked \target{carnations} and
  \target{dahlias}.)
\end{myexe}

In cases, when the sentiment relation is represented by a comparison,
you should also mark each of the compared objects as a separate
target.  Additionally, for the object which is being preferred in the
comparison, you also should set the \textit{preferred} attribute of
that target to \textit{true}:
\begin{myexe}
  Ich mag $<$\textbf{target
    preferred=``true''}$>$Domino-Eis$<$\textbf{/target}$>$ mehr als
  \target{Magnum}.

  (I like the $<$\textbf{target preferred=``true''}$>$Domino ice
  cream$<$\textbf{/target}$>$ more than \target{Magnum}.)
\end{myexe}

Possible attributes and attribute values for the \texttt{target} tag are
listed in Table \ref{tbl-target}:
\begin{center}
  \begin{tabular}{|l|c|p{\clmnwidth}|}\hline
    Attribute & Value & Value's Meaning\\\hline\label{tbl-target}

    & \textit{preferred} (default) & in comparisons, this value means that the
    respective target is being considered better than another compared object,
    e.g. \textit{\textbf{Die neue Frisur} passt ihr garantiert besser als die
      alte (\textbf{The new hairstyle} suits her definitely better than the
      old one)};\\\cline{2-3}

    \multirow{-2}{*}{preference} & \textit{dispreferred} & in
    comparisons, this value marks the target element which is being
    considered worse than its counterpart, e.g. \textit{Die zweite
      Saison von Breaking Bad war viel spannender als \textbf{die
        dritte} (The second season of Breaking Bad was much more
      exciting than \textbf{the third one})};\\\hline

    sentiment-ref & \textit{$\longrightarrow$\newline(directed edge)} &
    directed edge pointing from the target element to its respective sentiment
    span.  You need to draw this edge in two cases:
    \begin{itemize}
    \item if the target element is located in an intersection of two different
      sentiments (in this case, you should draw an edge from the target to the
      sentiment relation, which this target actually belongs to)

    \item when the target of a sentiment opinion is expressed outside the
      sentiment span
    \end{itemize}\\\hline

    anaph-ref & \textit{$\longrightarrow$\newline(directed edge)} & directed
    edge pointing from the target element expressed by a pronoun or pronominal
    adverb to its respective non-pronomial antecedent (in order to draw this
    edge, you will also need to mark the antecedent as target)\\\hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Source
\subsection{source}
The \texttt{source} tag is used to mark the immediate author(s) or
holder(s) of a sentiment opinion.  These are usually speakers or
writers of an assessment or persons whose assessing attitude is being
cited.  In cases, when the source is not explicitly present in the
message, it is implicitly assumed to be the author of the tweet.

An example of a \texttt{source} expression is the pronoun \textit{Sie
  (she)} in the following sentence:
\begin{myexe}
  \source{Sie} mag die neue Farbe nicht

  (\source{She} doesn't like the new color)
\end{myexe}

You should usually mark the source element after you have determined and
annotated the sentiment relation and its target.  The source should denote the
immediate author or holder of the determined sentiment opinion.  In case of
citations, only the person or institution, who originally expressed the
assessment, should be marked as \texttt{source} (see Example
\ref{ex:source-requote}).
\begin{myexe}
  Laut Staatsanwalt soll die \source{Angeklagte} sich missbilligend \"uber
  ihren Vorgesetzten ge\"au\ss{}ert haben.

  (According to the attorney, the \source{defendant} had made disapproving
  remarks about her boss.)\label{ex:source-requote}
\end{myexe}
The citing persons should not be considered as sources unless it is explicitly
stated that they also share the opinion which they cite.

A \texttt{source} tag can usually include:
\begin{itemize}
  \item Pronouns, e.g. \textit{\source{Er} genoss die Aussicht von seinem
    Balkon (\source{He } enjoyed the view from his balcony)}.  In this case,
    you should also draw an edge from the pronoun to its respective antecedent
    if this antecedent can be observed in the immediate context (see
    \texttt{anaph-ref} attribute in Table \ref{tbl-target});

  \item Noun phrases formed by nouns and their attributes,
    e.g. \textit{\source{Sein Onkel} regte sich \"uber die neuen Steuern auf
      (\source{His uncle} got worked up over the new taxes)};
\end{itemize}
In cases, when the source of a sentiment is expressed by multiple
coordinatively joined independent noun phrases, you should mark each such noun
phrase with a separate \texttt{source} tag, e.g.:
\begin{myexe}\label{ex:2source}
  \source{Ihr} und \source{ihrer Mutter} gef\"allt die neue Farbe
  nicht.\\ (\source{She} and \source{her mother} do not like the new
  color)
\end{myexe}

The attributes of the \texttt{source} tag are listed in Table
\ref{tbl-source}.  The meaning of these attributes is fully identical
to the meaning of the corresponding attributes for \texttt{target}s.
\begin{center}
  \begin{tabular}{|l|c|p{\clmnwidth}|}\hline
    Attribute & Value & Value's Meaning\\\hline\label{tbl-source}

    sentiment-ref & \textit{$\longrightarrow$\newline(directed edge)}
    & see Table \ref{tbl-target}\\\hline

    anaph-ref & \textit{$\longrightarrow$\newline(directed edge)} &
    see Table \ref{tbl-target}\\\hline
  \end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Emo-expression
\subsection{emo-expression}
With the \texttt{emo-expression} tag, you should mark words or
phrases, that have some polar subjective sense in their lexical
meaning.  An example of a possible emo-expression is given below:
\begin{myexe}
  Beim Aufr\"aumen des Zimmers haben wir einen \emoexpression{ekelhaften }
  Teller mit verschimmeltem Essen unter Bett gefunden.

  (When we cleaned the room, we found a \emoexpression{disgusting}
  plate of moldy food under the bed.)
\end{myexe}

You should however note, that, in contrast to the sources and targets, which
should only be annotated in the context of a sentiment relation, you should
always mark an emotional expression regardless of whether a sentiment is
present or not.

On the other hand, because of the lexical ambiguity, it can also sometimes
happen that the same word or expression, which appears to have an
emotional lexical sense in one context, will have a neutral meaning in
another.  Here, you should mark an element as an \texttt{emo-expression} only
in the former case, i.e. when its immediately activated sense in the given
text is polar (see the following examples):
\begin{myexe}
  Dieser Wein ist ein echtes \emoexpression{Juwel} in meiner Kollektion.

  This wine is a real \emoexpression{jewel} in my collection.
\end{myexe}
\begin{myexe}
  Koh-i-Noor ist das teuerste Juwel heutzutage.

  Koh-i-Noor is the most expensive jewel nowadays.
\end{myexe}
In the first example, the meaning of \textit{Juwel (jewel)} is metaphoric and
only this metaphoric sense has a polar connotation.  In the second example, the
meaning of \textit{Juwel (jewel)} is literal and does not have any appraising
aspect.

The \texttt{emo-expression}s are usually represented by:
\begin{itemize}
  \item nouns, e.g. \textit{Held (hero)}, \textit{Ideal (ideal)},
    \textit{Betr\"uger (fraudster)} etc.;

  \item adjectives or adverbs, e.g. \textit{sch\"on (nice)},
    \textit{zuverl\"assig (reliably)}, \textit{hinterh\"altig
      (devious)}, \textit{heimt\"uckisch (insidiously)} etc.;

  \item verbs, e.g. \textit{lieben (to love)}, \textit{bewundern (to
    admire)}, \textit{hassen (to hate)} etc.;

  \item idioms, e.g. \textit{auf die Nerven gehen (to get on the
    nerves)} etc.;

  \item smileys, e.g. :), :-(, \smiley{}, \frownie{} etc.;
\end{itemize}
This list is however neither complete nor limited.

The \texttt{emo-expression} elements can accept the following
attributes with the following possible values:
\begin{center}
  \begin{tabular}{|l|c|p{\clmnwidth}|}\hline\label{tbl-emo}
    Attribute & Value & Value's Meaning\\\hline
    %%%%%%%%%%%%%%%%%%%

    & \textit{positive} & given emotional
    expression has a positive appraising meaning, e.g. \textit{gut
      (good), verhimmeln (to ensky), Prachtkerl (corker)}
    etc.\\\cline{2-3}

    \multirow{-2}{*}{polarity} & \textit{negative\newline(default)} & given
    emotional expression has a negative polar sense, e.g. \textit{versauen (to
      botch up), rotzig (snotty), Dreckskerl (scum)} etc.\\\hline

    %%%%%%%%%%%%%%%%%%%

    & \textit{weak} & given emo-expression has a weak appraising sense,
    e.g. \textit{solala (so-so), nullachtf\"unfzehn (vanilla),
      durchschnittlich (mediocre)} etc.\\\cline{2-3}

    & \textit{medium\newline(default)} & given emo-expression has a regular
    appraising sense and middle stylistic expressivity, e.g. \textit{gut
      (good), schlecht (bad), robust (tough)} etc.\\\cline{2-3}

    \multirow{-3}{*}{intensity} & \textit{strong} & given emo-expression forms
    a very strong positive or negative assessment, e.g. \textit{allerbeste
      (bettermost), zum Kotzen (to make one puke), Kacke (cack)} etc.\\\hline

    %%%%%%%%%%%%%%%%%%%

    sentiment-ref & \textit{$\longrightarrow$\newline(directed edge)} & this
    attribute has the same meaning as its equivalents for targets and sources.
    In ambiguous usage cases, it should show to which sentiment the given
    emo-expression belongs.  Just like for targets, a usage case should be
    considered as ambiguous if this expression is located in the intersection
    of two sentiment spans or if it is expressed outside of any marked
    sentiment relation but has an immediate relation to it.  If an
    emo-expression does not belong to any sentiment or if it is unambigously
    embedded in just one sentiment relation, this attribute edge needs not to
    be drawn.\\\hline
  \end{tabular}
\end{center}

When determining the polarity of an emotional expression, you should keep in
mind the following two aspects:
\begin{itemize}
  \item The annotated polarity of an \texttt{emo-expression} should be its
    \textit{prior} (or sometimes also called \textit{lexical}) polarity.  This
    type of polarity is the polar sense of a word or expression irrespective
    of its possible contextual changers.  In simple words, this usually means
    that you should disregard any negations when determining and setting the
    value of the polarity attribute for an \texttt{emo-expression}. So, for
    example, in the sentence \textit{Es war keine gute Idee (It was not a good
      idea)}, the polarity of the \texttt{emo-expression} \textit{gute (good)}
    is still positive despite the fact that in the end, this polarity is
    negated;

  \item Some emo-expressions can have an assymmetric polarity for their
    arguments.  A typical example of such expression is the word
    \textit{vermissen (miss)}.  If we say that \textit{Sarah vermisst Andreas
      (Sarah misses Andreas)}, it usually means something negative for Sarah
    (because she experiences a discomfort feeling) but at the same time
    something positive for Andreas (because it is implied that Sarah would be
    happier if he joined her).  In these cases, if one of the arguments is the
    sentiment's target, you should determine the polarity of the
    emo-expression from the perspective of the target.  So, in this case, the
    target of the sentiment is \textit{Andreas} (it is his absence which is
    being assessed) and the polarity of the emo-expression \textit{vermissen
      (miss)} should therefore be set to positive.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Intensifier
\subsection{intensifier}
Intensifiers are elements which increase the stylistic expressivity and the
polar sense of an emotional expression.  An example of an identifier is the
word \textit{sehr (very)} in the following example:
\begin{myexe}
  wir suchen eine \intensifier{sehr} zuverl\"assige Polin als
  Haushaltshilfe.

  (we are looking for a \intensifier{very} reliable Polish woman as
  domestic help.)
\end{myexe}
An intensifier should always relate to some emo-expression and you should
always explicitly show this relation by drawing an attribute edge from the
intensifier to the respective expression whose sense is being intensified
(see the \texttt{emo-expression-ref} attribute in Table \ref{tbl-intens}).

Intensifiers are usually expressed by adverbs or adjectives like, for example,
\textit{sehr (very), sicherlich (certainly)} etc.  Though other ways of
expressing these elements are still possible.

The intensifier elements can have the following attributes with the
following possible values:
\begin{center}
  \begin{tabular}{|l|c|p{\clmnwidth}|}\hline\label{tbl-intens}

    & \textit{medium (default)} & the intensifier moderately increases
    the polar sense of the emotional expression, e.g. \textit{ziemlich
      (quite), recht (fairly)} etc.\\\cline{2-3}

    \multirow{-2}{*}{degree} & \textit{strong} & the intensifier
    strongly increases the polar sense and stylistic markedness of the
    emotional expression, e.g. \textit{sehr (very), super (super),
      stark (strongly)} etc.\\\hline

    %%%%%%

    emo-expression-ref & \textit{$\longrightarrow$\newline(directed
      edge)} & a directed edge pointing from the intensifier to the
    \texttt{emo-expression} whose meaning is being intensified\\\hline
  \end{tabular}
\end{center}

\subsection{diminisher}
With the \texttt{diminisher} tag, you should mark elements which decrease the
polar lexical sense of an emotional expression.  In Example
\ref{ex:diminisher}, you can see one possible usage of a diminisher:
\begin{myexe}
  \diminisher{Weniger} erfolgreiche Unternehmen verzichten auf externe
  Berater.\label{ex:diminisher}

  The \diminisher{less} successful companies do not use external
  consultants.
\end{myexe}
Just like intensifiers, diminishers should always relate to some emotional
expression and you should always explicitly show this relation by drawing the
attribute edge \texttt{emo-expression-ref} (see Table \ref{tbl-intens}).

Also like intensifieres, diminishers are usually expressed by adjectives or
adverbs, e.g. \textit{wenig (few), kaum (hardly), ein bisschen (little)}.
Though, other means of expressing these elements are also possible.

Diminishers have the same set of attributes as intensifiers with the only
difference in the interpretation of the \textit{degree} attribute:
\begin{center}
  \begin{tabular}{|l|c|p{\clmnwidth}|}\hline\label{tbl-dimin}

    & \textit{medium (default)} & given diminisher moderately
    decreases the polar sense of its respective emotional expression,
    e.g. \textit{wenig (few), bisschen (little)} etc.\\\cline{2-3}

    \multirow{-2}{*}{degree} & \textit{strong} & given diminisher
    strongly decreases the polar sense of the emotional expression,
    e.g. \textit{kaum (hardly)} etc.\\\hline

    emo-expression-ref & \textit{$->$\newline(directed edge)} &
    see Table \ref{tbl-intens}\\\hline
  \end{tabular}
\end{center}

\subsection{negation}
With the \texttt{negation} tag, you should mark elements, which turn the
polarity of a sentiment relation to its complete opposite.  For example, in
the next sentence, the negative article \textit{kein (not)} makes sentiment's
polarity negative, though, without this article, it would be positive:
\begin{myexe}
Einstein war \negation{kein} kluger Mensch!\label{ex:negation}

Einstein was \negation{not} a clever person!
\end{myexe}

The role and the meaning of the negations are closely related to that of the
diminishers.  In order to help you better differentiate between the two, we
have listed the most obvious differences between these classes:
\begin{itemize}
  \item\textit{Semantic differences}.  While diminishers still allow
    for some portion of the original meaning of an emo-expression to
    remain (\textit{a hardly understandable speech} is still
    understandable), the negations fully deny this meaning and change
    it to a completely antonymous one (\textit{a not understandable
      speech} is completely unintelligible);

  \item\textit{Part-of-speech differences}.  While diminishers are usually
    expressed by either adjectives or adverbs; negations, on the contrary, are
    commonly represented by either the negative article \textit{kein (no)} or
    the negation particle \textit{nicht (not)}, or complete clauses, e.g.
    \textit{Es ist sehr zweifelhaft, dass die neue Version von Windows besser
      wird (It is very doubtful that the new Windows version will be any
      better)}.

  %% \item\textit{Syntactic differences}.  While diminishers are usually
  %%   expressed by either adjectives or adverbs; negations, on the contrary,
  %%   are commonly represented by either the negative article \textit{kein
  %%   (no)} or the negation particle \textit{nicht (not)}, or even complete
  %%   clauses (\textit{Er ist nicht einverstanden, dass die neue Version von
  %%   Windows besser ist (He disagrees that the new Windows version is any
  %%   better)}).
\end{itemize}
All three elements -- diminishers, intensifiers, and negations -- should only
be annotated in the context of a sentiment relation.  If no sentiment is
present in the message, you should not annotate these elements either.

As we already mentioned, negations can usually be expressed by:
\begin{itemize}
  \item the negative article \textit{kein (no)}, e.g. \textit{Es war keine
    Vergn\"ugungsreise f\"ur australische Machthaber (It was no pleasure trip
    for some of Australia's most powerful leaders)};
  \item the negation particle \textit{nicht (not)}, e.g. \textit{Er hat seine
    Studienzeit ganz und gar nicht genossen (He did not enjoy his college days
    at all)};
  \item clauses, e.g. \textit{Es ist absolut unm\"oglich, dass aus ihm am Ende
    noch ein guter Mensch wird (It is absolutely impossible that he will
    eventually become a good man)}.
\end{itemize}
\textcolor{red}{The negation elements have no attributes.}

\section{Summary}\label{sec:summary}
To summarize, your task in this assignment is to find subjective appraising
opinions, that judge about some particular things or events from a personal
perspective.  You should annotate these assessments as \texttt{sentiment}s,
determine whether the judgements they represent are negative or positive, and
also decide how strong the judgement is.  Next, you should separately mark the
object being assessed as sentiment \texttt{target}.  The author of the opinion
(or the person whose appraising opinion is being cited) should be marked as
sentiment \texttt{source}.  Both, sources and targets should only be marked if
a sentiment relation is present in message.

Another important task is to annotate lexical expressions whose meaning
contains some polar assessing sense.  These elements are called
\texttt{emo-expression}s and they should always be annotated regardless of
whether a sentiment relation is present or not.  In case when an
emo-expression, however, gives rise to a sentiment and the meaning of the
expression is being intensified, diminished, or negated by some elements, you
should also annotate these modifying elements as \texttt{intensifier},
\texttt{diminisher}, and \texttt{negation}, correspondingly.

\section{FAQ}\label{sec:faq}
This section provides possible solutions to some difficult and controversial
annotation cases.

\begin{itemize}
\item\textbf{F: Should I annotate as sentiments cases like \textit{einer Sache
    zustimmen (to agree with sth.)}, \textit{sich f\"ur etwas entscheiden (to
    opt for sth.)} or \textit{etwas unterst\"utzen (to support sth.)}?}

  Well, these cases are a little bit tricky and, unfortunately, we will not be
  able to give you a clean and elegant solution for them.  We would rather say
  that both variants, i.e. annotating and not annotating sentences with these
  elements as sentiments, would be legitimate.  In fact, after all annotators
  finish their work, we are going to conduct a study on the inter-annotator
  agreement for such expressions and make our final decision based on the
  majority of choices made by different annotators.

\item\textbf{F: How would you annotate the following cases of comparisons?}
  \begin{itemize}
  \item\textbf{\textit{Seehofer hat die Gr\"unen ausgeschlossen , aber die
      Linke nicht (Seehofer has excluded the Greens, but not the Left)}};

    Without any further context, I cannot see any sentiment relation here.
    So, I would probably not annotate anything;

  \item\textbf{\textit{Lieber starke Mitte statt linker Rand (Better strong
      middle than left edge)}};

    Comparison with \textit{starke Mitte (strong middle)} as preferred target,
    and \textit{linker Rand (left edge)} as dispreferred one;

  \item\textbf{\textit{Die \#spd wird lieber mit den rechten von \#cdu , \#csu
      koalieren als mit der \#linke (The \#spd will better for a coalition
      with the rightists from \#cdu, \#csu than with the \#left)}};

    Comparison, with \textit{\#spd} as source, \textit{\#cdu} and
    \textit{\#csu} preferred targets, and \textit{\#linke (\#left)} as
    dispreferred one;

  \item\textbf{\textit{Die \#AfD + vereinigt mehr \"okonomische Kompetenz als
      alle etabl. Parteien + Bunde... (The \#AfD + combines more economic
      expertise than all established parties + federal...)}};

    Comparison, with \textit{\#AFD} as preferred target and
    \textit{established parties} and \textit{federal} as dispreferred ones;

  \item\textbf{\textit{Freiheit statt Bevormundung (Freedom instead of
      paternalism)}};

    Comparison, with \textit{Freiheit (freedom)} as preferred target, and
    \textit{Bevormundung (paternalism)} as dispreferred one;

  \item\textbf{\textit{Fettarme Milch hat mittlerweile mehr Prozent wie die
      FDP (lowfat milk has meanwhile more percents than the FDP)}};

    I'd rather say that this is sarcasm about the FDP.  Because we usually
    cannot compare a bottle of milk with a political party.  If we do so then
    usually with the purpose of tutting at this party;

  \item\textbf{\textit{Was ist der Unterschied zwischen einem Smart und der
      FDP ? Der Smart hat wenigstens 2 Sitze :) (What is the difference
      between a Smart and the FDP? The Smart has at least two seats)}};

    The same as the previous question -- sarcasm about the FDP;
  \end{itemize}

\item\textbf{F: How can I determine the intensity of a comparison?}

  A: Well, comparisons are, in fact, assessments too.  But in contrast to
  other sentiment relations, these assessments are relative rather than
  absolute.  It means that we do not claim that something is bad or good, but
  we just say that something is worse or better than something else.  The more
  vivid this difference between two compared objects is, the higher is the
  intensity of the comparison.  So, if we find \textit{this lousy Telekom
    waaaaaay less reliable than O2}, then the intensity should probably be
  higher than in \textit{Telekom has a less reliable connection than 02}.

\item\textbf{F: Should I annotate sentiments in recommendations?}

  A: Usually yes.  If this recommendation represents a personal opinion and is
  not a spam, and you can unambiguously determine the target of this opinion,
  then you should annotate sentiment here.

\item\textbf{F: Should I annotate sentiments in defenses?}

  A: Usually not.  If a soldier defends his position or a PhD student defends
  her thesis, it does not necessarily imply that he or she likes it.  The same
  is true in cases when someone defenses another person in a dispute.

\item\textbf{F: Should I annotate sentiments in questions?}

  A: Well, it primarily depends on the type of the question.  Here, you should
  distinguish two cases:
  \begin{itemize}
    \item If it is a \textit{yes-no-question}, which asks whether a particular
      sentiment opinion is true or not, you should not annotate any sentiment
      relation in such sentence.  For example, in \textit{Gef\"allt dir der
        neue Rock?  (Do you like the new skirt?)}, we do not know whether the
      asked person likes or dislikes the new skirt, thus we should not
      annotate a sentiment in this case;
    \item If it is a \textit{wh-question}, which asks about the reasons or
      some other aspects of a polar opinion but does not raise the truth of
      this opinion to question, the sentiment relation should be marked.  For
      example, in \textit{Warum hasst du deine Schwester? (Why do you hate
        your sister?)}, the fact that someone hates her sister is undoubtful
      and we should annotate sentiment in this sentence.
  \end{itemize}

\item\textbf{F: Should I mark sentiments in insults?}

  A: If you can locate the target, then yes.  For example, in \textit{Du bist
    ein Idiot! (You are an idiot!)}, \textit{Du (You)} is the target of a
  negative sentiment relation.  On the contrary, in cases like \textit{Idiot!
    (Idiot!)}, no target is detectable and therefore, according to our
  definition, no sentiment can exist.

\item\textbf{F: Is it possible that sources and targets are expressed by other
  means than the ones described in these guidelines?}

  A: Yes. These guidelines are in no way exhaustive, they should just give you
  a better sense of how a typical source or target usually might look like.

\item\textbf{F: It is said that we should disregard negations when determining
  the polarity of an emo-expression.  What about sentiments, shall we take
  into account negations there?}

  A: Yes.  The polarity of an emo-expression represents the polar sense of a
  single lexical item.  The polarity of a sentiment, on the contrary, should
  show the joint sense of the whole phrase, which forms a sentiment relation,
  so you should take into account negations there.

\item\textbf{F: How should I annotate chains of intensifiers/diminishers --
  each separately or the whole chain with one tag?}

  A: In case of multiple intensifiers/diminishers, each element should be
  tagged separately, e.g. \textit{Du bist die '\intensifier{aller}
    \intensifier{aller}' Beste! (You are the \intensifier{very}
    \intensifier{very} best!)}

%% \item F: Dem Tweet aus politics1 ``Wo ist der \#Jubel von \#CDU \#CSU
%%   \& \#FDP \"uber den Tod der Mieterin nach \#Zwangsr\"aumung?'' habe
%%   ich als sarkastisch und mit target=''Tod der Mieterin nach
%%   \#Zwangsr\"aumung'' annotiert. Wei\ss aber nicht, ob das so richtig
%%   ist.

%% A: Mhm, aus meiner Sicht wird hier eher \"uber die CDU/CSU und FDP
%% gespottet, man w\"urde also annehmen, dass diese Parteien sich \"uber
%% den Tod der Mieterin freuen w\"urden, was nat\"urlich dem Image einer
%% Partei eher schaden w\"urde.  Ich w\"urde demzufolge den ganzen Satz
%% als negatives Sentiment mit den Parteien als Targets und dem Attribut
%% sarcasm gesetzt auf true annotieren.
\end{itemize}
\end{document}

\subsection{sentiment}
\begin{itemize}
\item I have difficulties determining the overall sentiment of a tweet.  Is
  there a common modus operandi?

Please keep in mind that sentiments should always be annotated
target-oriented. Look at this tweet: \textit{Ich hab Maria ja schon so
  vermisst und so , ne ? (I kind of did miss Maria, right?)}. Here you
might think the tweet has a negative sentiment, because the author
missed Maria which is not a good thing. But we are always only
concerned about the sentiment relation between the source and the
target! And here it is a positive sentiment, because the author missed
Maria because he likes her.

\item How to annotate sarcasm?

If the apparent form of a sentiment is positive (and therefore should
be annotated as a positive sentiment) but you can detect sarcasm
please annotate the polarity as negative and additionally set
\texttt{sarcasm} to \texttt{true} in the attribute window. This
applies to the sentiment tag as well as to emo-expressions.

\item Should subjunctives be annotated?

Yes. Have a look on the following tweet: \textit{W\"are toll wenn das
  der n\"achste " Call of Duty " - Teil sein w\"urde , liebe
  @GameStar. (Dear @gamestar, it would be great if this would be the
  sequel of "Call of Duty".)}. Although this is a subjunctive
sentiment and is not "reality" right now you should annotate this,
too.

\item Should multiple sentiment layers be annotated?

Yes. For example in the tweet \textit{72 j\"ahriger Leser regt sich
  bei mir \"uber YouTube und Gema auf :) (72 year old reader troubles
  over YouTube and Gema at me :) )} you can find two layers of
sentiments. The first layer is the fact that the reader troubles over
YouTube and Gema. The second layer is the fact that the author of the
tweet is amused about the first fact. So here you have to annotate two
different sentiments. The first is negative and has "YouTube" and
"Gema" as target (with "troubles over" as emo-expression). The second
is positive and has "72 jähriger Leser" as target and "mir" as
source. Do not forget to use the \texttt{sentiment\_ref} attribute to
annotate which sentiment refers to which target.
\end{itemize}

\subsection{source}
\begin{itemize}
\item I can find the same source two times in one tweet. What to do?

Look at this tweet: \texttt{Ich hab \"Ubrigens ne 1 in Englisch.  Find
  ich gut ! (By the way, I got an A- in English. I Like it!)}  Here
the "Ich" in the first sentence needs not to be annotated. It is
sufficient to annotate just the "Ich" in the second sentence. Since
there is no anaphora present you will not need to mark an anaphoric
antecedent as well. \\ Here you should annotate both sentences as one
sentiment with "ne 1- in Englisch" as target, the "I" in the second
sentence as source and "gut" as emo-expression.
\end{itemize}

\subsection{target}

\subsection{emo-expression}
\begin{itemize}
\item What about English emo-expressions?

Although we annotate German tweets, it is not unusual to find English
emo-expressions. For example: \textit{DSDS ist fancy ! (DSDS is fancy
  !)}.  These expressions should be annotated too, if they are common
in German.

\item Should intejections be marked as emo-expressions?

Sollen Interjektionen (oh, aha, achso, OMG etc.) auch als
emo-expressions annotiert werden? Ich habe das n\"amlich nicht
gemacht, weil sie in so gut wie allen F\"allen nicht als wirklich
positiv bzw. negativ zu identifizieren waren. Wie
z.B. \textit{Lichgestaaaaaaaaaaalt !!!!!! in deren Schatten ich mich
  drehe uuuooooooh oooooh uuuuuooohhhh oooooohhhhh xD} oder
\textit{ooh du hasts gut , h\"att auch gern urlaub ...}. Ich finde
Interjektionen auch grenzwertig, da sie unter gewissen Umst\"anden
auch als intensifier dienen k\"onnen, z.B. oh wie gut, oh ja
etc. . Falls sie doch annotiert werden sollen, bitte Bescheid sagen
und/oder eine entsprechende Anmerkung in den Guidelines machen. W\"are
gut explizit zu schreiben ob sie annotiert werden sollen oder nicht.

\item Do we need sarcasm attribute for emo-expressions?
\end{itemize}

\subsection{intensifier}
\subsection{diminisher}
\subsection{negation}
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Examples}
Please DO NOT mark as sentiments following cases:
\begin{itemize}
\item Statements describing some emotional states for which no
  target can be derived or found (e.g. \textit{Ich f\"uhle mich so
    traurig :(} - \textit{I am feeling so blue :(});
\item Statements describing some objective facts, even if possible
  consequences of these facts can be assumed to have negative
  influence on the author (e.g. \textit{Wenn ein Floh einen Menschen
    bei\ss{}t und ihn mit erbrochenem Blut infiziert, werden die
    Pestbakterien ins Gewebe \"ubertragen.} - \textit{When a flea
    bites a human and contaminates the wound with regurgitated
    blood, the plague carrying bacteria are passed into the
    tissue.});
\item Interrogative clauses in case they ask whether some polar
  opinion is true or not. (e.g. \textit{Findet die Mehrheit
    der Bev\"olkerung CDU toll?} - \textit{Does the majority
    of population consider CDU great?});
\end{itemize}

\subsection{negation}
\texttt{negation}s are lexical or syntactic elements that reverse the
primary polar meaning of emo-expressions to the opposite so that the
overall polarity of the whole sentiment is different to the polarity
of emo-expressions belonging to it. A typical example of negation is
\textit{nicht} in sentences like \textit{Ein guter Schritt war diese
  Entscheidung \textbf{nicht}.} (\textit{This decision was
  \textbf{not} a good move.}).

\vspace{0.5cm}
You SHOULD only mark negative elements that have an impact on the
sentiment polarity. Negating elements having no such impact
shold not be marked. Negation elements are usually represented by:
\begin{itemize}
\item Negation particle \textit{nicht} (\textit{not}),
  e.g. \textit{Klug ist dieser Hund sicherlich \textbf{nicht}}
  (\textit{This dog is certainly \textbf{not} clever});

\item Negative article \textit{kein} e.g. \textit{Er war
  \textbf{kein} Vorbild f\"ur seine Kinder} (\textit{He was
  \textbf{not} a good role model for his children});

\item Indefinite pronouns like \textit{niemand}, \textit{keiner}
  etc., e.g. \textit{\textbf{Niemand} hielt ihn f\"ur einen
    ehrlichen Menschen.} (\textit{\textbf{Nobody} considered him an
    honest man});

\item Any lexical or idiomatic unit in case they turn the sentiment
  polarity to the opposite, e.g. \textit{Ich \textbf{zweifle}, dass
    das neue iPhone ein besseres Display hat.} (\textit{I
    \textbf{doubt} the new iPhone has a better display}).
\end{itemize}

DO NOT mark elements as negations that have no effect on the sentiment
polarity. For example, in the sentence \textit{Ich mag Leute, die nicht
  nur an sich selbst denken.} (\textit{I like people who not only care
  about themselves.}) \textit{nicht} (\textit{not}) should \underline{not} be
marked as negation since it does not change the positive polarity
expressed by \textit{m\"ogen} (\textit{like}).
\vspace{0.5cm}

Negations only have one possible attribute, namely
\textit{sentiment-ref} that is a directed edge pointing from a negation
to the sentiment it belongs to. You should only draw this edge in cases multiple sentiment 
relations overlap and it is not obvious to which of these sentiments the negation belongs to,
i.e. in cases the negation is used for both sentiment spans.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{intensifier}
\texttt{intensifier}s are elements that increase the polar
meaning of emotional expressions. Intensifiers are usually expressed
by adjectives or adverbs like \textit{sehr} (\textit{very}),
\textit{ziemlich} (\textit{rather}) and the like. In intensifier-chains 
\textit{(e.g. sehr, sehr gut)} please annotate each intensifier separately.
As mentioned above you should annotate every single emo-expression you can 
find regardless of a corresponding sentiment span. 
Please do not forget to annotate the corresponding intensifier if there is one. \newline

Intensifiers have the following attributes and values: \newline

\begin{tabular}{|l|c|p{\clmnwidth}|}\hline
  \multirow{2}{*}{degree} & \textit{1 (default)} & intensifier 
  slightly increases polarity of emotional
  expression, e.g. \textit{ziemlich},
  \textit{recht} etc.\\\cline{2-3}

  & \textit{2} & intensifier strongly increases polarity  of
  emotional expression, e.g. \textit{sehr}, \textit{super},
  \textit{stark} etc.\\\hline

  %%%%%%

  sentiment-ref & \textit{$->$\newline(directed edge)} & Directed
  edge pointing from the intensifier to the sentiment it belongs
  to. By analogy to negations, you should only draw this edge in cases
  multiple sentiment relations overlap and it is not obvious
  to which of these sentiments the intensifier belongs to,
  i.e. in cases when the intensifier is used for both
  sentiment spans.\\\hline
\end{tabular}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{diminisher}
\texttt{diminisher}s are the counterpart to \texttt{intensifier}s.
These are elements that decrease the polar meaning of emotional
expressions.  Like intensifiers diminishers are usually expressed by
adjectives or adverbs. Typical examples of such adverbs are
\textit{wenig}, \textit{kaum}, \textit{ein bisschen} etc..

Diminishers have the same attributes as intensifiers. The only
difference are the values for the attribute. Instead of positive
values you have negative ones.

Here is a table of diminisher's attributes:

\begin{tabular}{|l|c|p{\clmnwidth}|}\hline

  \multirow{2}{*}{degree} & \textit{weak (default)} & diminisher
  slightly decreases polarity of emotional expression,
  e.g. \textit{wenig}, \textit{bisschen} etc.\\\cline{2-3}

  & \textit{strong} & diminisher strongly decreases polarity of
  emotional expression, e.g. \textit{kaum} etc.\\\hline


  sentiment-ref & \textit{$->$\newline(directed edge)} & Directed
  edge pointing from the diminisher to the sentiment it belongs
  to. You should only draw this edge in cases
  multiple sentiment relations overlap and it is not obvious
  to which of these sentiments the diminisher belongs to,
  i.e. in cases when the diminisher is used for both
  sentiment spans.\\\hline

\end{tabular}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{MMAX Techniques}
In this section the most important techniques for annotating our corpus in MMAX will be described.

\subsection{Turn on Auto-Save and Auto-apply}
After setting up MMAX load a .mmax file as described in the beginning
of this document. You will be asked to validate the
annotations. Please always do that.  Before you start the annotation
it is recommended to turn on "Auto-Save" and "Auto-apply". Go to
\texttt{File -> Auto-Save -> Every} and select a time value.  Then
change to the window for the attributes (the window in the upper left
corner of your screen) and go to \texttt{Settings} and check the
\texttt{Auto-apply} box.  Your choice will be confirmed by a red
"Auto-apply is ON" at the bottom of the window. \newline

\subsection{Annotate Markables and Attributes}
The window in the middle of your screen now should show the tweets. To annotate a markable choose a word, phrase, clause or sentence by clicking the left mouse-button and mark the respective span while holding the mouse-button. After releasing the left mouse-button a menu will pop up showing the list of markables. Left-click on the corresponding markable and the span will turn to the respective color of the markable. Then change to the attribute window in the upper left corner, click on the tab for the markable you just annotated and set the value(s) for the attributes. (In case you did not turn on "Auto-apply" press "Apply".)

\subsection{Deleting Markables}
Just do a \underline{right-click} on a colored markable and left-click \texttt{Delete this markable}.

\subsection{Annoating Discontinuous Markables}
Sometimes you will find discontinuous emo-expressions. The most prominent example are German particle verbs (e.g. weh tun). To annotate them as one single element first annotate one element (it will turn to the respective markable color) as usual, left-click to highlight it and then mark the other element. A menu will pop up saying "Add to this markable". Left-click on that.

\subsection{Annotating anaphref}
Imaginge you have a sentence like: \textit{Peter ist schlau und gut aussehen tut er auch noch. (Peter is clever and he looks good, too.)}. Here you should annotate \textit{er} as anaphref to \textit{Peter}. To do so first annotate \textit{Peter} and \textit{er} as single targets. Then left-click on \textit{Peter} to highlight the markable and then do a \underline{right-click} on \textit{er}. A menu will pop up saying \texttt{"Mark as anaphric antecendent of target"}. Left-click on that.

\subsection{Annotating sentiment\_ref}
The same technique as above applies. First mark a sentiment span as usual. Second annotate the other markable (whatever it may be) as usual. Left-click on the first markable to highlight it then \underline{right-click} on the second markable and in the pop-up menu left-click on "Mark sentiment which this span belongs to".

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{FAQ}

\underline{Things you SHOULD mark as sentiments:}
\begin{itemize}
  \item Phrases, clauses, sentences and statements expressing a polar opinion
    (e.g. \textit{Der neue Film \"uber Superman war knorke!} -
    \textit{The new superman movie was fantastic});
  \item Interrogative clauses containing a polar opinion
    and do not call this opinion into question but rather ask for its
    reasons or other related aspects (e.g. \textit{Warum findet die
      Mehrheit der Bev\"olkerung CDU so toll?} - \textit{Why does the
      majority of the population consider CDU great?});
  \item Exclamations expressing support or disapproval of
    something (e.g. \textit{Der Sommer ist da. Super!} -
    \textit{Summer has come. Great!}).
  \item Comparisons expressing some preference of one subject/object over another 
      (e.g. \textit{Canon EOS 550d macht bessere Aufnahmen als 600d} - \textit{Canon EOS 550d takes
      better pictures than 600d}).
\end{itemize}


\underline{Things you SHOULD NOT mark as sentiments:}
\begin{itemize}
  \item Statements that do not have an annotatable target:
      (e.g. \textit{F\"uhle mich so traurig :(} - \textit{Feeling so blue :(});
  \item Statements describing some objective facts, even if possible
    consequences of these facts can be assumed to have negative
    influence on the author (e.g. \textit{Wenn ein Floh einen Menschen
      bei\ss{}t und ihn mit erbrochenem Blut infiziert, werden die
      Pestbakterien ins Gewebe \"ubertragen.} - \textit{When a flea
      bites a human and contaminates the wound with regurgitated
      blood, the plague carrying bacteria are passed into the
      tissue.});
  \item Interrogative clauses in case they ask whether some polar
    opinion is true or not. (e.g. \textit{Findet die Mehrheit
      der Bev\"olkerung CDU toll?} - \textit{Does the majority
      of the population consider CDU great?});
\end{itemize}

\underline{Things you SHOULD mark as sources:}
\begin{itemize}
  \item Original author(s) of polar opinions, i.e. persons,
    groups or officials who express sympathy or antipathy for
    something or somebody e.g. \textit{\textbf{Michael} meinte, das w\"are die
      beste L\"osung.}  (\textit{\textbf{Michael} thought, it would be the
      best solution.});
\end{itemize}


\underline{Things you SHOULD NOT mark as sources:}
\begin{itemize}
  \item Persons how are neutrally citing someone else's opinion,
    e.g. in the sentence \textit{Nach Tatjanas Worten war Michael sehr
      dar\"uber ver\"argert.} (\textit{According to Tatjana, Michael
      was very angry about that.}) you should only mark
    \textit{Michael} as source and not \textit{Tatjana}. Please note, that in case an author
    supports or contradicts someone else's opinion, two sentiment
    relations should be made - one with the original author as source
    and one with the supporter/opponent of the opinion as the other
    source.\footnote{In that case sentiments will overlap, and
      you also should mark these sources with \textit{sentiment-ref}-tag.}
\end{itemize}

\underline{Things you SHOULD mark as emo-expressions:}
\begin{itemize}
  \item Adjectives and adverbs bearing polar attitudes \textit{Peter
    hatte \textbf{bessere} Noten in der Schule als sein Bruder}
    (\textit{Peter had \textbf{better} grades at school than his
    brother.});

  \item Verbs expressing attitude of a speaker to target,
    e.g. \textit{Mir \textbf{gefiel} die neue House-Staffel}
    (\textit{I \textbf{liked} the new House series});

  \item Idiomatic expression including support verbs
    e.g. \textit{\textbf{Zum Teufel} soll die neue Regierung
      \textbf{gehen}} (\textit{The new government should \textbf{go to
        hell}}).

  \item Smileys in case they really express an emotional attitude and
    are not used for politeness or without any particular meaning
    e.g. \textit{Gleich in Braunschweig mit Kameraden treffen
        \textbf{:)}} (\textit{Will soon meet friends in Braunschweig
        \textbf{:)}}).
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
